He-Man: A domain-specific language for lightweight concurrency

### 2011-12-13 report/report.tex
Finished backend section. Report finished?

\begin{figure}[ht]\label{fig:backend}
\centering
\bnf{
\i{Thread}
\p{(\i{thread\_label}, [(\i{Var}, \i{Type})])}
\i{Block}
\p{(\i{label}, \i{thread\_label}, [\i{Stmt}], \i{Tail})}
\i{Tail}
\p{\b{Goto} \i{label}}
\a{\b{GotoWait} \i{label}}
\a{\b{If} \i{Expr} \i{Tail} \i{Tail}}
\a{\b{Exit}}
\i{Stmt}
\p{\b{Decl} (\i{Var}, \i{Type}) \i{Expr}}
\a{\b{Assign} \i{Expr} \i{Expr}}
\a{\b{Exp} \i{Expr}}
\a{\b{Spawn} \i{label} [((\i{Var}, \i{Type}), \i{Expr})]}
}
\caption{Back-end grammar.}
\end{figure}

The most important task for our back-end is to convert our linear front-end code
into a collection of states appropriate for an event loop. This occurs in our
\t{flattenPrgm} pass, which passes through the program in reverse order,

is to compile C,

It's important that the only blocking operations happen in the main loop

We separate

Threads are separated and replaced by labels

@@

From this point, the path to C code is fairly straightforward. For each thread,
we generate a \t{struct} \t{thread$n$\_t} declaration containing a \t{thread\_t}
struct with generic thread control information, followed by all thread-local
variables. Each block is translated to a function taking as argument a
\t{thread\_t} passed in from the main loop. It casts this to the \t{struct} type
for its thread in order to access thread-local variables.

>>>>

\begin{figure}[ht]
\centering
\bnf{
\i{Thread}
\p{(\i{thread\_label}, [(\i{Var}, \i{Type})])}
\i{Block}
\p{(\i{label}, \i{thread\_label}, [\i{Stmt}], \i{Tail})}
\i{Tail}
\p{\b{Goto} \i{label}}
\a{\b{GotoWait} \i{label}}
\a{\b{If} \i{Expr} \i{Tail} \i{Tail}}
\a{\b{Exit}}
\i{Stmt}
\p{\b{Decl} (\i{Var}, \i{Type}) \i{Expr}}
\a{\b{Assign} \i{Expr} \i{Expr}}
\a{\b{Exp} \i{Expr}}
\a{\b{Spawn} \i{label} [((\i{Var}, \i{Type}), \i{Expr})]}
}
\caption{Back-end grammar.}
\label{fig:back-syntax}
\end{figure}

The next several passes convert our front-end grammar to a back-end grammar,
shown in Figure \ref{fig:back-syntax}, which closely resembles the features of
the C code that we subsequently generate. Most significantly, a program now
consists of a list of \i{Block}s of code, each of which belongs to a \i{Thread},
and a list of \i{Thread}s. \i{Expr}s remain unchanged from the front-end, while
all control flow is delegated to the \i{Tail} position of a \i{Block} of code.

As the grammar makes clear, the most important task for our back-end is to
convert our linear front-end code into a collection of states appropriate for an
event loop. This occurs in our \t{flattenPrgm} pass, which passes through the
program in reverse order with a \t{RWS} monad, converting it to uniquely-named
blocks whose only control-flow consists of tail-position jumps to other block
labels. Here, the \t{State} maintains the next block number, the \t{Writer}
collects lists of \i{Block}s and \i{Thread}s, and the \t{Reader} maintains the
current thread number (defined as the first block in the current thread).

As an illustrative example, consider encountering an \b{If} statement in our
backwards pass, where we have access to the translation of the later statements,
as well as the earlier statements. Morally, we would like to translate this to a
conditional jump to the translation of either the first or second arm, each of
which is terminated by a jump to the later statements. (Note that
\t{flattenPrgm} leaves all optimization to the next pass.)

We package up the later statements in a new block, and then translate the
preceding statements, terminated by a tail-position conditional \b{If}. We also
translate each arm of the original \b{If}, terminated by a tail-position
\b{Goto} to the block containing the later statements. Importantly, all three of
these recursive translations may generate new blocks, which will be collected
automatically by the \t{Writer} monad.

Because blocking I/O may only happen in the main loop, we also terminate a block
after any blocking I/O call, by inserting a \b{GotoWait} to the block containing
the later statements. \b{GotoWait} is identical to \b{Goto} except to signal to
later passes that the thread \emph{must} yield to the main loop at this
location, and thus the goto must not be optimized away.

@@

From this point, the path to C code is fairly straightforward. Concretely, we
transform our back-end grammar to the AST provided by the Language.C package,
which we then output (as C source) to a file. For each thread, we generate a
\t{struct} \t{thread$n$\_t} declaration containing a \t{thread\_t} struct with
generic thread control information, followed by all thread-local variables. Each
block is translated to a function taking as argument a \t{thread\_t} passed in
from the main loop. It casts this to the \t{struct} type for its thread in order
to access thread-local variables.
